Static Discharge of Higher-Order Interposition

### 2013-10-02 pldi2014/paper.tex
Cleaned up semantics and other additions.

impersonators and chaperones

The \scheme{chaperone-procedure} procedure takes a procedure \scheme{p} to be chaperoned as its first argument and a procedure \scheme{w} to wrap \scheme{p} as its second argument.
The argument and result arities of \scheme{p} and \scheme{w} must be compatible.
This means that the argument arity of \scheme{w} must be the same as
%\footnote{Racket procedures can include both mandatory and optional keyword arguments. When considering these, it is necessary only that the set of mandatory keyword arguments of \scheme{p} include those of \scheme{w} and the set of all keyword arguments of \scheme{p} be included in.}
The result of a \scheme{chaperone-procedure} call is a procedure which, upon invocation, will deliver the arguments to \scheme{w}.
\begin{enumerate}
\item The $n$ arguments are delivered to \scheme{w}.
\item If \scheme{w} returns $n$ results, the results are 
\end{enumerate}

@@

An impersonator wraps a value and intercepts operations on them, potentially redirecting them.
In the case of procedures--first-class values in Racket--impersonators intercept arguments and can alter or replace them before passing them to the procedure they wrap.
As they pass these arguments, they have the opportunity to register for interest in intercepting the results of the application by including an extra argument: a callback procedure which will be applied to the results of the inner procedure.

@@

@@

@@

%Chaperone values are branded with the label of their creation site. In any given program, there are a finite number of these creation sites and labels. In fact, it is these creation sites that we are concerned with. The output of an analysis is a map from labels to transformations (no-op, promotion, discharge). Once we apply these transformations, we recover optimization opportunities as data dependency decreases.

%We call a wrapper function $g$ an \emph{originator} of a chaperone $c_0$ if $c_0$ is created in within the extent of $g$. Chaperone values are also branded with their parent chaperon

%Definition: call a creation site good0 if all of the chaperone values which flow from it are well-behaved.

@@

%\begin{theorem}[Selective Chaperone Discharge Theorem]
%A good and nice chaperone installment site can soundly be replaced by its first argument %(where ``soundly'' means ``with no change to the meaning of the program'').
%\end{theorem}

@@

%Strickland presents a chaperone erasure theorem which states that if a program with chaperones evaluates to a value, the same program with chaperones erased--that is, with all chaperone installment sites replaced with the first argument--will evaluate to the same value. The purpose of this theorem seems to be to make a statement about the nature of chaperones--that, in a well-behaved program, they do not change meaning. Our theorems give the conditions of when a given chaperone installment site does not change the meaning of the program.

@@

%In the Racket implementation, the \scheme{chaperone-procedure} primitive is access most often through a wrapper function that handles keyword procedures.
%Our static analysis determines the transformation of a \scheme{chaperone-procedure} application site to be sound only if no contract installed there will ever be blamed.
%This means that the violation of a single contract installed there prevents the promotion or discharge of \emph{any} of the other contracts installed there.
%In order for the analysis to be effective, we need more granularity.
%We preprocess programs to inline certain definitions according to human judgement.

We define \chapcalc, or \emph{chaperone calculus}, an extension of the A-normalized $\lambda$-calculus.
With the introduction of chaperones and impersonators (and the incusion of primitive functions), we can no longer count on the operator of an application to be a closure.
Thus, we will treat these three categories of values as classes of \emph{operators} and will refer to them collectively as such.

The syntactic class $\ae$ signifies \emph{atomic expressions} and includes $\lambda$-terms, variables, and the simple values of booleans, integers, and primitive functions--expressions which will always produce a value, as opposed to raising an error or diverging.

Both \scheme{chaperone-operator} and \scheme{impersonate-operator} exist as the tags of syntactic forms instead of first-class procedure values as they exist in Racket.
This is solely (is it?) to simplify the presentation of the analysis.
In the actual analysis, we consider all call sites and add a condition for the sound discharge of a call site that \scheme{chaperone-operator} is the only values to flow to the operator position.

The second subform of a \scheme{chaperone-operator} or \scheme{impersonate-operator} remains a proper expression to facilitate some of the semantics.

$e:=\lete{x\,\dots}{c}{e}$ where $\vx=\langle x_1,\dots x_n\rangle$

$\letrece{x\,\dots}{c}{e}$ where $\vx=\langle x_1,\dots x_n\rangle$

values = < > + - not car cdr null? pair? cons null integer? boolean? raise

We define the semantics of \chapcalc in terms of an abstract machine similar to Felleisen's CESK machine[cite that] in order to ease the transition to a pushdown system.
In difference, our machine is defined in terms of multiple state variants.
\begin{itemize}
\item An \emph{expression} state, tagged $\Sigma_{e}$, corresponds to a traditional CESK state.
(The order of the components is reversed, however, to aid factoring out common components from among the classes.)
\item An \emph{apply} state, tagged $\Sigma_{a}$, represents a point of application and dispatches on the operator type (closure, impersonator, or primitive).
\item A \emph{value} state, tagged $\Sigma_{v}$, represents a point of return and dispatches on the type of the top continuation frame.
If the continuation is empty, the constituent values are the results of the program.
\item An \emph{error} state, tagged $\Sigma_!$, represents the arisal of an error and, as the result of the program, may include blame information.
\item Finally, we define a set of \emph{dummy} states which splits a transition which would otherwise pop and push a frame simulaneously into a sequence of two transitions which performs each separately.
(This eases further the transition to a pushdown system.)
\end{itemize}

\newcommand{\ks}[0]{\kappa s}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\Sigma_e(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\Sigma_v(#1,#2,#3)}
\newcommand{\sapp}[4]{\Sigma_a(#1,#2,#3,#4)}
\newcommand{\scha}[6]{\Sigma_d(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdum}[6]{\Sigma_d(#1,#2,#3,#4,#5,#6)}
\newcommand{\serr}[0]{\mathrm{error}}
\newcommand{\sbla}[1]{\mathrm{blame}(#1)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\red}[2]{#1\longrightarrow #2}

\newcommand{\bind}[4]{\mathrm{bind}(#1,#2,#3,#4)}

\newcommand{\clos}[2]{(#1,#2)}
\newcommand{\chap}[2]{\mathrm{chap}(#1,#2)}

\newcommand{\letk}[3]{\mathrm{let}_\kappa(#1,#2,#3)}
\newcommand{\letreck}[3]{\mathrm{letrec}_\kappa(#1,#2,#3)}
\newcommand{\ifk}[3]{\mathrm{if}_\kappa(#1,#2,#3)}
\newcommand{\ifd}[3]{\delta_\mathrm{if}(#1,#2,#3)}
\newcommand{\andk}[2]{\mathrm{and}_\kappa(#1,#2)}
\newcommand{\ork}[2]{\mathrm{or}_\kappa(#1,#2)}

\subsection{$\chapcalc$ Reduction}

We give the pertinent reduction rules for each form in the grammar.

\noindent
$\red{\sexp{\ks}{\sigma}{\rho}{\appe{\ae_f}{\ae_1\,\dots\,\ae_n}}}{\sapp{\ks}{\sigma}{f}{\langle v_1,\dots,v_n\rangle}}$

\noindent
$\red{\sexp{\ks}{\sigma}{\rho}{\chae{\ae_f}{e_w}}}{\sexp{\chak{f}::\ks}{\sigma}{\rho}{e_w}}$

\noindent
where $f=\A{\sigma}{\rho}{\ae_f}$ and $\mathrm{operator?}(f)$.

\noindent
$\red{\sexp{\ks}{\sigma}{\rho}{\chae{\ae_f}{e_w}}}{\sbla{\ell}}$

\noindent
where $f=\A{\sigma}{\rho}{\ae_f}$ and $\mathrm{operator?}(f)$ does not hold.

\noindent
$\red{\sval{\chak{f}::\ks}{\sigma}{\vv}}{\sval{\ks}{\sigma}{\langle\chap{\alpha_f}{\alpha_w}\rangle}}$

\noindent
where $\alpha_f=\mathrm{alloc}(\sigma,f)$ and $\alpha_w=\mathrm{alloc}(\sigma,w)$
if $\vv=\langle w\rangle$ for some value $w$, $\mathrm{operator?}(w)$, and $\mathrm{arity}(w)=\mathrm{arity}(f)$.

\noindent
$\red{\sval{\chak{f}::\ks}{\sigma}{\vv}}{\sbla{\ell}}$

\noindent
if $\vv\ne\langle w\rangle$ for any value $w$, $\mathrm{operator?}(w)$ does not hold, or $\mathrm{arity}(w)\ne\mathrm{arity}(f)$.

$\red{\sexp{\ks}{\sigma}{\rho}{\impe{e_f}{e_w}}}{}$

\noindent
$\red{\sexp{\ks}{\sigma}{\rho}{\lete{x_1\,\dots\, x_n}{e_0}{e_1}}}{\sexp{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\rho}{e_0}}$

$\red{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e_1}}$ where $(\sigma',\rho')=\bind{\sigma}{\rho}{\vx}{\vv}$ if $|\vx|=|\vv|$

$\red{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\serr}$ if $|\vx|\ne|\vv|$

$\red{\sexp{\ks}{\sigma}{\rho}{\letrece{\vx}{e_0}{e_1}}}{\sexp{\letreck{\rho}{\vx}{e_1}::\ks}{\sigma}{\rho}{e_0}}$

$\red{\sexp{\ks}{\sigma}{\rho}{\ife{e_t}{e_c}{e_a}}}{\sexp{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\rho}{e_t}}$

$\red{\sval{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma}{\rho}{e_c}}$
if $\vv=\langle v_0\rangle$ for some $v_0\ne\false$

$\red{\sval{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma}{\rho}{e_a}}$
if $\vv=\langle v_0\rangle$ for $v_0=\false$

$\red{\sval{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\vv}}{\serr}$
if $\vv\ne\langle v_0\rangle$ for any $v_0$

A closure operator proceeds straightforwardly, extending its environment and evaluating its body or raising an error for an arity mismatch.

$\red{\sapp{\ks}{\sigma}{\clos{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e}}$
where $\vx=\langle x_1,\dots,x_n\rangle$ and $(\sigma',\rho')=\bind{\sigma}{\rho}{\vx}{\vv}$ if $|\vx|=|\vv|$

$\red{\sapp{\ks}{\sigma}{(\lambda\mathbf{x}.e,\rho)}{\vv}}{\dynbla{\ks}}$
if $|\vx|\ne|\vv|$

A chaperone as an operator recurs on its wrapper and pushes a frame to complete the interposition when its wrapper returns.

$\red{\sapp{\ks}{\sigma}{\mathrm{chaperone}(f,w)}{\vv}}{\sapp{\chacwk{f}{\vv}::\ks}{\sigma}{w}{\vv}}$

If the number of results obtained is the same as the number of arguments given, the interposition is complete and the function can be evaluated in tail position with respect to the chaperone application.

$\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sapp{\ks}{\sigma}{f}{\vvp}}$
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

$\red{\sval{\chacfk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sdum{\ks}{\sigma}{f}{\vvp}{\ell}{w}}$
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

$\red{\sdum{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\sapp{\chacfk{w}::\ks}{\sigma}{f}{\vvp}}$

$\red{\sval{\chacfk{w}::\ks}{\sigma}{\vv}}{\sdum{\ks}{\sigma}{w}{\vv}{\ell}{what}}$

$\red{\sdum{\ks}{\sigma}{w}{\vv}{\ell}{what}}{\sapp{\chacrk{\vv}::\ks}{\sigma}{w}{\vv}}$

$\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sval{\ks}{\sigma}{\vvp}}$
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

>>>>

%impersonators and chaperones

%The \scheme{chaperone-procedure} procedure takes a procedure \scheme{p} to be chaperoned as its first argument and a procedure \scheme{w} to wrap \scheme{p} as its second argument.
%The argument and result arities of \scheme{p} and \scheme{w} must be compatible.
%This means that the argument arity of \scheme{w} must be the same as
%\footnote{Racket procedures can include both mandatory and optional keyword arguments. When considering these, it is necessary only that the set of mandatory keyword arguments of \scheme{p} include those of \scheme{w} and the set of all keyword arguments of \scheme{p} be included in.}
%The result of a \scheme{chaperone-procedure} call is a procedure which, upon invocation, will deliver the arguments to \scheme{w}.
%\begin{enumerate}
%\item The $n$ arguments are delivered to \scheme{w}.
%\item If \scheme{w} returns $n$ results, the results are 
%\end{enumerate}

\setspecialsymbol{LAMBDA}{$\lambda$}
\setspecialsymbol{GE}{$\ge$}
\setspecialsymbol{EPS}{$\epsilon$}

@@

An impersonator wraps a value and intercepts operations on them, potentially redirecting them.
In the case of procedures--first-class values in Racket--impersonators intercept arguments and can alter or replace them before passing them to the procedure they wrap.
As they pass these arguments, they have the opportunity to register to intercept the results of the application by including an extra argument: a callback procedure which will be applied to the results of the inner procedure.

@@

For instance, we might create a factorial function which dynamically enforces the type of its argument and result with
\begin{schemedisplay}
(chaperone-procedure
 (letrec ([fact (LAMBDA (n)
                  (if (= n 0)
                      1
                      (* n (fact (- n 1)))))])
   fact)
 (LAMBDA (n)
   (if (and (integer? n) (GE n 0))
       (values (LAMBDA (n)
                 (if (and (integer? n) (GE n 1))
                     n
                     (raise positive error ...)))
               n)
       (raise negative error ...))))
\end{schemedisplay}
Keeping the results wrapper enclosed lexically in the arguments wrapper allows it to capture arguments for later comparison with results.
This allows dependent contracts to be expressed with chaperones.
For example,
\begin{schemedisplay}
(chaperone-procedure
 sqrt
 (LAMBDA (x)
   (if (GE x 0)
       (values (LAMBDA (y)
                 (if (< (abs (- (* y y) x)) EPS)
                     y
                     (raise positive error ...)))
               x)
       (raise negative error ...))))
\end{schemedisplay}
Here the specification for the \scheme{sqrt} procedure is encoded in the chaperoning procedure.

@@

The monitoring chaperones perform on the procedures they wrap takes time and space, but this does not represent all of the run-time cost associated with them.
Chaperones themselves (and impersonators alike) are monitored to meet a minimum standard of well-behavedness.
For instance, the runtime ensures that each guard function returns an appropriate number of results, which number depends on the number of arguments supplied at application.
In addition, arity compatibility of the positive guard must be checked at application time as it is undecidable in general to determine it before the guard itself and the values to which it is to be applied are in hand.
Where the standard of well-behavedness of chaperones deviates from that of impersonators is in the restriction that guard results must be the same as the arguments given or chaperones of them.

@@

% the spirit of this is to motivate a calculus

Both \scheme{chaperone-operator} and \scheme{impersonate-operator} exist as the tags of syntactic forms instead of first-class procedure values as they exist in Racket.
Providing them as first-class values increases the expressive power [cite Felleisen] of Racket and, in general, requires a more powerful analysis.
In practice, the values are most often referenced directly by name in the operator position.
We treat them as syntactic forms to simplify the presentation of the analysis, but discuss our strategy to integrate first-class uses of them in Section ?.
%In the actual analysis, we consider all call sites and add a condition for the sound discharge of a call site that \scheme{chaperone-operator} is the only values to flow to the operator position.

We define \chapcalc, or \emph{chaperone calculus}, an extension of the A-normalized $\lambda$-calculus.

\subsection{\chapcalc\ Syntax}

The syntactic class $\ae$ signifies \emph{atomic expressions} and includes $\lambda$-terms, variables, and the simple values of booleans, integers, and primitive functions. These expressions will never produce an error or diverge but will always produce a value.

The second subform of a \scheme{chaperone-operator} or \scheme{impersonate-operator} remains a proper expression to facilitate some of the blame semantics.

$e:=\lete{x_1\,\dots x_n}{c}{e}$

$\letrece{x_1\,\dots x_n}{c}{e}$

values = < > + - not integer? boolean? raise

\newcommand{\ks}[0]{\kappa s}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\Sigma_e(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\Sigma_v(#1,#2,#3)}
\newcommand{\sapp}[4]{\Sigma_a(#1,#2,#3,#4)}
\newcommand{\scha}[6]{\Sigma_d(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdchaz}[6]{\Sigma_{chap_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdchao}[5]{\Sigma_{chap_1}(#1,#2,#3,#4,#5)}
\newcommand{\serr}[0]{\mathrm{error}}
\newcommand{\sbla}[2]{\mathrm{blame}_{#1}(#2)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\red}[2]{\begin{align*}& #1\\\rr\, & #2\end{align*}}

\newcommand{\bind}[4]{\mathrm{bind}(#1,#2,#3,#4)}
\newcommand{\prebind}[2]{\mathrm{pre\mhyphen bind}(#1,#2)}
\newcommand{\recbind}[4]{\mathrm{rec\mhyphen bind}(#1,#2,#3,#4)}

\newcommand{\clo}[2]{(#1,#2)}
\newcommand{\cha}[2]{\mathrm{chaperone}(\ell,#1,#2)}
\newcommand{\imp}[2]{\mathrm{impersonator}(#1,#2)}

\newcommand{\impk}[1]{\mathrm{imp}_\kappa(\ell,#1)}
\newcommand{\impcwk}[2]{\mathrm{imp\mhyphen neg\mhyphen ults}_\kappa(\ell,#1,#2)}
\newcommand{\impcfk}[1]{\mathrm{imp\mhyphen fun\mhyphen ults}_\kappa(\ell,#1)}
\newcommand{\impcrk}[1]{\mathrm{imp\mhyphen pos\mhyphen ults}_\kappa(\ell,#1)}

\newcommand{\letk}[3]{\mathrm{let}_\kappa(#1,#2,#3)}
\newcommand{\letreck}[3]{\mathrm{letrec}_\kappa(#1,#2,#3)}

\newcommand{\rr}{\longrightarrow}
\newcommand{\rrs}{\longrightarrow^{*}}

\subsection{$\chapcalc$ Semantics}

We define the semantics of \chapcalc\ in terms of an abstract machine similar to Felleisen's CESK machine[cite that] in order to ease the transition to a pushdown system.

With the introduction of chaperones and impersonators (and the incusion of primitive functions), we can no longer count on the operator of an application to be a closure.
Thus, we will treat these three categories of values as classes of \emph{operators} and will refer to them collectively as such.

In difference, our machine is defined in terms of multiple state variants.
\begin{itemize}
\item An \emph{expression} state, tagged $\Sigma_{e}$, corresponds to a traditional CESK state.
(The order of the components is reversed, however, to aid factoring out common components from among the classes.)
\item An \emph{apply} state, tagged $\Sigma_{a}$, represents a point of application and dispatches on the operator type (closure, impersonator, or primitive).
\item A \emph{value} state, tagged $\Sigma_{v}$, represents a point of return and dispatches on the type of the top continuation frame.
If the continuation is empty, the constituent values are the results of the program.
\item An \emph{error} state, tagged $\Sigma_!$, represents the arisal of an error and, as the result of the program, may include blame information.
\item Finally, we define a set of \emph{dummy} states which splits a transition which would otherwise pop and push a frame simulaneously into a sequence of two transitions which performs each separately.
(This eases further the transition to a pushdown system.)
\end{itemize}

A program is injected into a machine state with the function $\mathcal{I} : e\rightarrow\Sigma$ defined as $\mathcal{I}(p)=\sexp{\langle\rangle}{\perp}{\perp}{p}$.

The reduction relation is defined first in terms of the class of machine state.

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\appe{\ae_f}{\ae_1\,\dots\,\ae_n}}}{\sapp{\ks}{\sigma}{f}{\langle v_1,\dots,v_n\rangle}}

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\chae{\ae_f}{e_w}}}{\sexp{\chak{f}::\ks}{\sigma}{\rho}{e_w}}

\noindent
where $f=\A{\sigma}{\rho}{\ae_f}$.

\noindent
\red{\sval{\chak{f}::\ks}{\sigma}{\vv}}{\sval{\ks}{\sigma}{\langle\cha{f}{w}\rangle}}

\noindent
%where $\alpha_f=\mathrm{alloc}(\sigma,f)$ and $\alpha_w=\mathrm{alloc}(\sigma,w)$
if $\vv=\langle w\rangle$ for some value $w$, $\mathrm{operator?}(f)$, $\mathrm{operator?}(w)$, and $\mathrm{arity}(w)=\mathrm{arity}(f)$.

\noindent
\red{\sval{\chak{f}::\ks}{\sigma}{\vv}}{\sbla{2,8,9}{\ell}}

\noindent
if $\vv\ne\langle w\rangle$ for any value $w$, $\mathrm{operator?}(f)$ does not hold, $\mathrm{operator?}(w)$ does not hold, or $\mathrm{arity}(w)\ne\mathrm{arity}(f)$.

%\red{\sexp{\ks}{\sigma}{\rho}{\impe{e_f}{e_w}}}{}

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\impe{\ae_f}{e_w}}}{\sexp{\impk{f}::\ks}{\sigma}{\rho}{e_w}}

\noindent
where $f=\A{\sigma}{\rho}{\ae_f}$.

\noindent
\red{\sval{\impk{f}::\ks}{\sigma}{\vv}}{\sval{\ks}{\sigma}{\langle\imp{f}{w}\rangle}}

\noindent
if $\vv=\langle w\rangle$ for some value $w$, $\mathrm{operator?}(f)$, $\mathrm{operator?}(w)$, and $\mathrm{arity}(w)=\mathrm{arity}(f)$.

\noindent
\red{\sval{\impk{f}::\ks}{\sigma}{\vv}}{\dynbla{\ks}}

\noindent
if $\vv\ne\langle w\rangle$ for any value $w$, $\mathrm{operator?}(f)$ does not hold, $\mathrm{operator?}(w)$ does not hold, or $\mathrm{arity}(w)\ne\mathrm{arity}(f)$.

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\lete{x_1\,\dots\, x_n}{e_0}{e_1}}}{\sexp{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\rho}{e_0}}

\red{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e_1}}

\noindent
where $(\sigma',\rho')=\bind{\sigma}{\rho}{\vx}{\vv}$ if $|\vx|=|\vv|$

\red{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\dynbla{\ks}} if $|\vx|\ne|\vv|$

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\letrece{x_1\,\dots\, x_n}{e_0}{e_1}}}{\sexp{\letk{\rho'}{\vx}{e_1}::\ks}{\sigma}{\rho}{e_0}}

\noindent
where $\vx=\langle x_1,\dots,x_n\rangle$ and $\rho'=\prebind{\rho}{\vx}$.

\red{\sval{\letreck{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma'}{\rho}{e_1}}

\noindent
where $\sigma'=\recbind{\sigma}{\rho}{\vx}{\vv}$ if $|\vx|=|\vv|$

\red{\sval{\letreck{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}{\dynbla{\ks}} if $|\vx|\ne|\vv|$

\red{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}{\sexp{\ks}{\sigma}{\rho}{e_c}}
if $v_t\ne\false$
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

\red{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}{\sexp{\ks}{\sigma}{\rho}{e_a}}
if $v_t=\false$
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

%$\red{\sval{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma}{\rho}{e_c}}$
%if $\vv=\langle v_0\rangle$ for some $v_0\ne\false$

%$\red{\sval{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\vv}}{\sexp{\ks}{\sigma}{\rho}{e_a}}$
%if $\vv=\langle v_0\rangle$ for $v_0=\false$

%$\red{\sval{\ifk{\rho}{e_c}{e_a}::\ks}{\sigma}{\vv}}{\serr}$
%if $\vv\ne\langle v_0\rangle$ for any $v_0$

The application of a closure operator proceeds straightforwardly by extending its environment and evaluating its body.

\red{\sapp{\ks}{\sigma}{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}{\sexp{\ks}{\sigma'}{\rho'}{e}}
where $\vx=\langle x_1,\dots,x_n\rangle$ and $(\sigma',\rho')=\bind{\sigma}{\rho}{\vx}{\vv}$ if $|\vx|=|\vv|$

If the number of values applied is incompatible with the arity of the closure, blame is assigned dynamically.

\red{\sapp{\ks}{\sigma}{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}{\dynbla{\ks}}
if $|\vx|\ne|\vv|$ where $\vx=\langle x_1,\dots,x_n\rangle$

Chaperone application recurs on its guard and augments the context to complete the interposition when its guard returns.
The enlarged context is annotated with the label of the applied chaperone (which itself is the label of its installment site).

\red{\sapp{\ks}{\sigma}{\cha{f}{w}}{\vv}}{\sapp{\chacwk{f}{\vv}::\ks}{\sigma}{w}{\vv}}

If the number of results obtained is the same as the number of arguments given, the interposition ceases and the function can be evaluated in tail position with respect to the chaperone application.

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sapp{\ks}{\sigma}{f}{\vvp}}
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sbla{1}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$ does not hold.

--

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sdchaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}
if $|\vv|=|\vvp|$, $\mathrm{operator?}(w)$, and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sbla{3}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{operator?}(w)$ does not hold.

\red{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}{\sbla{1}{\ell}}
if $|\vv|=|\vvp|$ and $\mathrm{operator?}(w)$ but $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$ does not hold.

--

\red{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}{\sbla{4}{\ell}}
if $|\vv|\ne|\vvp|$ and $|\vv|+1\ne|\vvp|$

\red{\sdchaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\sapp{\chacfk{w}::\ks}{\sigma}{f}{\vvp}}

\red{\sval{\chacfk{w}::\ks}{\sigma}{\vv}}{\sdchao{\ks}{\sigma}{w}{\vv}{\ell}}

\red{\sdchao{\ks}{\sigma}{w}{\vv}{\ell}}{\app{\chacrk{\vv}::\ks}{\sigma}{w}{\vv}}

\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sval{\ks}{\sigma}{\vvp}}
if $|\vv|=|\vvp|$ and $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$

\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sbla{7}{\ell}}
if $|\vv|=|\vvp|$ but $\mathrm{chaperone\mhyphen of?}(\vvp,\vv)$ does not hold

\red{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}{\sbla{6}{\ell}}
if $|\vv|\ne|\vvp|$

%$\mathrm{apply}(\kappa s,\sigma,\mathrm{chap}(f,w),\mathbf{v})=\mathrm{apply}(\mathrm{chap\mhyphen ults}_\kappa(f,\mathbf{v})::\kappa s,\sigma,w,\mathbf{v})$

\subsection{Expectation that many dynamic checks are unnecessary}

Programs that can result in an error do not implement their specification; in other words, they are incorrect.
The purpose of blame assignment is to facilitate the act of correcting the program.
We expect a program to result in a blame assignment only rarely.
We have similar expections for a program to result in divergence, and that specifically at the hand of a chaperoned value.
In fact, we expect correct programs never to exhibit such behavior, and while chaperones serve the reality of incorrect programs, the frequency of their complaints diminishes in the long run as the programmer applies their feedback.
The nature of the programming cycle indicates that there may be some chaperones which, situated within a particular program, do not influence the behavior of the program and also that the proportion of chaperoned values in such a position may increase as the program is maintained.

\subsection{Intuition behind discharge}

If neither a chaperone nor its installation will raise an error or diverge for any possible execution of the program, it is a candidate for \emph{discharge}, or the replacement of its installation site with its first argument.
It is only a candidate because it may install a chaperone in the course of its monitoring and that chaperone may not be so well-behaved.
Discharging the original chaperone will prevent a meaningful chaperone from ever being created which makes the discharge unsound.
Then, intuitively, a chaperone can be discharged if it will not raise an error or diverge for any possible execution \emph{and} any chaperones it may install can be discharged.

However, multiple chaperone installations may occur at a given chaperone installation site during program execution, so we cannot, in all circumstances, selectively prevent the creation of a single chaperone.
Instead, we consider the discharge of installment sites themselves.
The conditions to discharge a given site are stronger than that of a chaperone installed at that site: instead of demonstrating that this single chaperone satisfies a property, we must demonstrate that every chaperone created at that site satisfies that property.
This constraint at first appears fatal.
Indeed, in Racket, users typically access chaperone installation through a library wrapper so all chaperones in the entire program share the same installation site.
This, in turn, means that a given chaperone can be safely discharged only if \emph{all} chaperones installed by any execution of the program can be safely discharged.
Fortunately, with access to the whole program, we may inline definitions which allows us to multiply the number of installment sites in the program.

%In such cases, we may not be able to discharge only a particular chaperone.
%We therefore strengthen the discharge condition that

\subsection{Intuition behind demotion}

Much of the run-time overhead of chaperones is shared by impersonators.
Nevertheless, the enforcement of certain invariants specific to chaperones is somewhat costly, specifically, the enforcement that chaperone guards at most install chaperones on their arguments.
The conditions to \emph{demote} the set of chaperones installed at a particular site to impersonators (by replacing the \scheme{chaperone-operator} installment site with \scheme{impersonate-operator}) are strictly weaker than to discharge the site entirely: we need to demonstrate only that the chaperone guards do not arbitrarily modify their arguments for it to be a candidate for demotion.
If one of this site's chaperones is installed on the argument of a chaperone guard, demoting this site to merely install an impersonator will make an otherwise well-behaved chaperone not so.
This observation provides us with the intuition behind sound site demotion: a chaperone installment may be demoted if neither of its guards do more than install chaperones on their arguments for all program executions and the chaperone which installs it, if any, can and is demoted. 
We can lift the conditions to chaperone installment sites so that a chaperone installment site can be demoted if all chaperones installed there in any program execution can be demoted and the installment site of any possible installing chaperone can and is demoted.

\subsection{More formal notion of discharge and demotion}

\subsection{Discharge soundness theorem}

If $p\rrs v$

Chaperone values are branded with the label of their creation site. In any given program, there are a finite number of these creation sites and labels. In fact, it is these creation sites that we are concerned with. The output of an analysis is a map from labels to transformations (no-op, promotion, discharge). Once we apply these transformations, we recover optimization opportunities as data dependency decreases.

We call a wrapper function $g$ an \emph{originator} of a chaperone $c_0$ if $c_0$ is created in within the extent of $g$. Chaperone values are also branded with their parent chaperon

Definition: call a creation site good0 if all of the chaperone values which flow from it are well-behaved.

@@

\begin{theorem}[Selective Chaperone Discharge Theorem]
A good and nice chaperone installment site can soundly be replaced by its first argument (where ``soundly'' means ``with no change to the meaning of the program'').
\end{theorem}

@@

Strickland presents a chaperone erasure theorem which states that if a program with chaperones evaluates to a value, the same program with chaperones erased--that is, with all chaperone installment sites replaced with the first argument--will evaluate to the same value. The purpose of this theorem seems to be to make a statement about the nature of chaperones--that, in a well-behaved program, they do not change meaning. Our theorems give the conditions of when a given chaperone installment site does not change the meaning of the program.

@@

In the Racket implementation, the \scheme{chaperone-procedure} primitive is access most often through a wrapper function that handles keyword procedures.
Our static analysis determines the transformation of a \scheme{chaperone-procedure} application site to be sound only if no contract installed there will ever be blamed.
This means that the violation of a single contract installed there prevents the promotion or discharge of \emph{any} of the other contracts installed there.
In order for the analysis to be effective, we need more granularity.
We preprocess programs to inline certain definitions according to human judgement.

\subsection{Demotion soundness theorem}
